---
title: "Austin Crime Analyis Markdown File"
author: "Rudy Duran"
date: "10/10/2020"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading the libraries


```{r}
library(tidyverse)
library(caret)
library(car)
library(ggplot2)
library(lubridate)
library(tree)
library(ggmap)
library(RColorBrewer)
library(randomForest)
library(sp)
library(expss)
library(corrplot)
```



# Loading the datasets
```{r}
austin_crime<- read.csv("C:\\Users\\duran\\Desktop\\Data Analytics Applications\\Final Project\\Austin Crime Dataset.csv")
```


# Checking the datatypes
```{r}
str(austin_crime)
```



# Getting rid of missing values
```{r}
austin_crimedata <- austin_crime
summary(austin_crimedata)
sum(is.na(austin_crimedata))
cleandata <- na.omit(austin_crimedata)
```



# Checking NAs are not there
```{r}
sum(is.na(cleandata))
```
# Getting rid of '' in the data
```{r}
cleandata = subset(cleandata, cleandata$Highest.Offense.Description != '')
cleandata = subset(cleandata, cleandata$Family.Violence != '')
cleandata = subset(cleandata, cleandata$Clearance.Status != '')
cleandata = subset(cleandata, cleandata$Occurred.Date.Time != '')
cleandata = subset(cleandata, cleandata$Occurred.Date != '')
cleandata = subset(cleandata, cleandata$Report.Date.Time != '')
cleandata = subset(cleandata, cleandata$Report.Date != '')
cleandata = subset(cleandata, cleandata$Location.Type != '')
cleandata = subset(cleandata, cleandata$Address != '')
cleandata = subset(cleandata, cleandata$APD.Sector != '')
cleandata = subset(cleandata, cleandata$APD.District != '')
cleandata = subset(cleandata, cleandata$PRA != '')
cleandata = subset(cleandata, cleandata$Clearance.Date != '')
cleandata = subset(cleandata, cleandata$UCR.Category != '')
cleandata = subset(cleandata, cleandata$Category.Description != '')
cleandata = subset(cleandata, cleandata$Location != '')
```


# Making a date out of Occurred.Date

```{r}
cleandata <- cleandata %>%
            mutate(Occurred_New_Date = mdy(Occurred.Date))
```



# Making a date out of Report.Date
```{r}
cleandata <- cleandata %>%
            mutate(Report_New_Date = mdy(Report.Date))
```




# Making a date out of Clearance Date
```{r}
cleandata <- cleandata %>%
            mutate(Clearance_New_Date = mdy(Clearance.Date))
```


# Getting year and month from occurred date


```{r}
cleandata <- cleandata %>%
            mutate(Occurred_Year = year(Occurred_New_Date),
                  Occurred_Month = month(Occurred_New_Date))
```



# Transforming Occurred.Date.Time column into a timestamp

```{r}
cleandata <- cleandata %>%
             mutate(Occurred_Time = mdy_hms(Occurred.Date.Time))
```



# Transforming Report.Date.Time into a timestamp\
```{r}
cleandata <- cleandata %>%
             mutate(Report_Time = mdy_hms(Report.Date.Time))
```


# Extracting the necessary time elements from Report_Time
```{r}
cleandata <- cleandata %>% mutate(Report_Year = year(Report_Time), 
                    Report_Month = month(Report_Time, label = TRUE), 
                    Report_Day = wday(Report_Time, label = TRUE))
```


# Extracting the necessary time elements from Occurred_Time
```{r}
cleandata <- cleandata %>% mutate(Occurred_Year = year(Occurred_Time), 
                    Occurred_Month = month(Occurred_Time, label = TRUE
                                           ), 
                    Occurred_Day = wday(Occurred_Time, label = TRUE))
```


# Subsetting the data for a specific timeframe

```{r}
cleandata <- subset(cleandata, Occurred_Year >= 2015)
```



# Factoring variables
```{r}
cleandata$Highest.Offense.Description <- as.factor(cleandata$Highest.Offense.Description)
cleandata$Family.Violence <- as.factor(cleandata$Family.Violence)
cleandata$APD.Sector <- as.factor(cleandata$APD.Sector)
cleandata$APD.District <- as.factor(cleandata$APD.District)
cleandata$PRA <- as.factor(cleandata$PRA)
cleandata$Location.Type <- as.factor(cleandata$Location.Type)
cleandata$Clearance.Status <- as.factor(cleandata$Clearance.Status)
cleandata$UCR.Category <- as.factor(cleandata$UCR.Category)
cleandata$Category.Description <- as.factor(cleandata$Category.Description)
cleandata$Zip.Code <- as.factor(cleandata$Zip.Code)
```



# Eliminating variables that are not needed

```{r}
cleandata = subset(cleandata, select = -c(Occurred.Date.Time, Occurred.Date, Occurred.Time, Report.Date.Time,
                                          Report.Date, Report.Time, Address,Location, Clearance.Date, 
                                           PRA, X.coordinate, Y.coordinate, Council.District, APD.Sector, APD.District,
                                          Census.Tract, UCR.Category ) )
```





# Making dummy day variables


```{r}

cleandata$dummymonday = ifelse(cleandata$Occurred_Day == "Mon", 1, 0)
cleandata$dummytuesday = ifelse(cleandata$Occurred_Day == "Tue", 1, 0)
cleandata$dummywednesday= ifelse(cleandata$Occurred_Day == "Wed", 1, 0)
cleandata$dummythursday = ifelse(cleandata$Occurred_Day == "Thu", 1, 0)
cleandata$dummyfriday = ifelse(cleandata$Occurred_Day == "Fri", 1, 0)
cleandata$dummysaturday = ifelse(cleandata$Occurred_Day == "Sat", 1, 0)
```


# Making Monthly dummy variables
```{r}
cleandata$dummyJanuary = ifelse(cleandata$Occurred_Month== "Jan", 1, 0)
cleandata$dummyFebruary = ifelse(cleandata$Occurred_Month == "Feb", 1, 0)
cleandata$dummyMarch= ifelse(cleandata$Occurred_Month == "Mar", 1, 0)
cleandata$dummyApril = ifelse(cleandata$Occurred_Month == "Apr", 1, 0)
cleandata$dummyMay = ifelse(cleandata$Occurred_Month == "May", 1, 0)
cleandata$dummyJune = ifelse(cleandata$Occurred_Month == "Jun", 1, 0)
cleandata$dummyJuly = ifelse(cleandata$Occurred_Month == "Jul", 1, 0)
cleandata$dummyAugust = ifelse(cleandata$Occurred_Month == "Aug", 1, 0)
cleandata$dummySeptember= ifelse(cleandata$Occurred_Month == "Sep", 1, 0)
cleandata$dummyOctober = ifelse(cleandata$Occurred_Month == "Oct", 1, 0)
cleandata$dummyNovember = ifelse(cleandata$Occurred_Month == "Nov", 1, 0)
```


# Checking last time on new dataset

```{r}
str(cleandata)
```


# Train/Test Splitting the data
```{r}
set.seed(1)
inTrain <- createDataPartition(cleandata$Family.Violence, p = 0.7, list = FALSE)
cleandata_train <- cleandata[inTrain,]
cleandata_test <- cleandata[-inTrain,]
```






# Logistic Regression
```{r}
model2 <- glm(Family.Violence~ Location.Type + dummymonday + dummytuesday + dummywednesday + dummythursday + dummyfriday + dummyJune  + dummyNovember , data = cleandata_train, family = "binomial")
summary(model2)
```





```{r}
# Running probabilities
cleandata_test$PredProb = predict.glm(model2, newdata = cleandata_test, type = 'response')
cleandata_test$PredChoice = ifelse(cleandata_test$PredProb >= 0.07, 1, 0)
cleandata_test$PredChoice = ifelse(cleandata_test$PredChoice == 1,'Y','N')
caret::confusionMatrix(as.factor(cleandata_test$Family.Violence), as.factor(cleandata_test$PredChoice))
```


# Random Forest on everything



```{r}
set.seed(1)
rf.training = randomForest(Family.Violence~., data = cleandata_train, importance = TRUE, ntree = 100)
```



# Checking variable importance plot
```{r}
varImpPlot(rf.training)
```





# Running Random Forest on specific variables

```{r}
set.seed(1)
rf.training = randomForest(Family.Violence~Location.Type + dummymonday + dummytuesday + dummywednesday + dummythursday + dummyfriday + dummyJune  + dummyNovember, data = cleandata_train, importance = TRUE, ntree = 100)
```

# Testing predictions on test dataset

```{r}
rf.pred = predict(rf.training, cleandata_test)
caret::confusionMatrix(rf.pred, cleandata_test$Family.Violence)
```

# Line Graphs- Question 2

Subsetting Data



```{r}
cleandata_occurred2017theft <- subset(cleandata, Occurred_Year == 2017 & Category.Description == "Theft")
cleandata_occurred2017murder <- subset(cleandata, Occurred_Year == 2017 & Category.Description == "Murder")
cleandata_occurred2017robbery <- subset(cleandata, Occurred_Year == 2017 & Category.Description == "Robbery")
cleandata_occurred2017burglary <- subset(cleandata, Occurred_Year == 2017 & Category.Description == "Burglary")
cleandata_occurred2017autotheft <- subset(cleandata, Occurred_Year == 2017 & Category.Description == "Auto Theft")
cleandata_occurred2017aggrassault <- subset(cleandata, Occurred_Year == 2017 & Category.Description == "Aggravated Assault")
```



# Running line plot on theft
```{r}
cleandata_occurred2017theft %>%
ggplot(aes(Occurred_Time)) +
  geom_freqpoly(binwidth = 86400) + 
  labs(title = "Austin Theft Occurrences- 2017",
                          x = NULL,
                          y = 'Number of Occurrences') + 
  theme_classic() +
    theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```


# Running line plot on murder
```{r}
cleandata_occurred2017murder %>%
ggplot(aes(Occurred_Time)) +
  geom_freqpoly(binwidth = 86400) + 
  labs(title = "Austin Murder Occurrences- 2017",
                          x = NULL,
                          y = 'Number of Occurrences') + 
  theme_classic() +
    theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```



# Running line plot on robbery
```{r}
cleandata_occurred2017robbery %>%
ggplot(aes(Occurred_Time)) +
  geom_freqpoly(binwidth = 86400) + 
  labs(title = "Austin Robbery Occurrences- 2017",
                          x = NULL,
                          y = 'Number of Occurrences') + 
  theme_classic() +
    theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Running line plot on burglary

```{r}
cleandata_occurred2017burglary %>%
ggplot(aes(Occurred_Time)) +
  geom_freqpoly(binwidth = 86400) + 
  labs(title = "Austin Burglary Occurrences- 2017",
                          x = NULL,
                          y = 'Number of Occurrences') + 
  theme_classic() +
    theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```


# Running line plot on auto theft
```{r}
cleandata_occurred2017autotheft %>%
ggplot(aes(Occurred_Time)) +
  geom_freqpoly(binwidth = 86400) + 
  labs(title = "Austin Auto Theft Occurrences- 2017",
                          x = NULL,
                          y = 'Number of Occurrences') + 
  theme_classic() +
    theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```


# Running line plot on aggravated assault
```{r}
cleandata_occurred2017aggrassault %>%
ggplot(aes(Occurred_Time)) +
  geom_freqpoly(binwidth = 86400) + 
  labs(title = "Austin Aggravated Assault Occurrences- 2017",
                          x = NULL,
                          y = 'Number of Occurrences') + 
  theme_classic() +
    theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```





# Getting the count for category descriptions
```{r}
Offense_count <- cleandata %>% 
              group_by(Highest.Offense.Description) %>% 
              summarize(count = n()) %>% 
              arrange(count)
print(Offense_count)
```

# Getting the top 10 offenses
```{r}
Top_10_Offense <- Offense_count %>%
                     top_n(10, count) %>%
                  arrange(desc(count))
print(Top_10_Offense)
```



# Getting the top 10 percentages
```{r}
Top_10_Offense <- Top_10_Offense %>%
                     mutate(pct = count/ sum(count))
print(Top_10_Offense)
```


# Getting Bottom 10


```{r}
Bottom_10_Offense <- Offense_count %>%
                     top_n(-10, count) %>%
                    arrange(desc(count)) %>%
                    mutate(pct = count/sum(count))

print(Bottom_10_Offense)
```


# Top 10 Offenses 2015-2019 Bar Graph

```{r}
ggplot(Top_10_Offense) + 
      geom_bar(mapping = aes(x = reorder(Highest.Offense.Description,+pct),y = pct, fill = Highest.Offense.Description), stat = 'identity') +
      coord_flip() + 
      labs(title = "Top 10 Austin Reported Incidents 2015-2019",
                          x = NULL,
                          y = NULL) + 
  theme_classic() +
    theme(legend.position = "none")+
  scale_y_continuous(labels = scales::percent_format())
```

# Bottom 10 Offenses 2015-2019 Bar Graph


```{r}
ggplot(Bottom_10_Offense) + 
      geom_bar(mapping = aes(x = reorder(Highest.Offense.Description,+pct),y = pct,fill = Highest.Offense.Description),   stat = 'identity') +
      coord_flip() + labs(title = "Bottom 10 Austin Reported Incidents 2015-2019",
                          x = NULL,
                          y = NULL) + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```

# Getting counts for locations
```{r}
Location_count <- cleandata %>% 
              group_by(Location.Type) %>% 
              summarize(count = n()) %>% 
              arrange(count)
print(Location_count)
```

# Getting counts for zip codes
```{r}
zip_count <- cleandata %>% 
              group_by(Zip.Code) %>% 
              summarize(count = n()) %>% 
              arrange(count)
print(zip_count)
```


# Getting percentages for zip codes
```{r}
zip_count <- zip_count%>%
                     mutate(pct = count/ sum(count))
print(zip_count)
```


# Austin Incidents Zip Code

Note: Could not recreate bar graph where all the bars were a specific type of blue as shown in the paper and presentation. 

```{r}
ggplot(zip_count) + 
      geom_bar(mapping = aes(x = reorder(Zip.Code,+count),y = count, fill = Zip.Code), stat = 'identity') +
      coord_flip() + 
      labs(title = "Incidents by Zip Code 2015-2019",
                        x = NULL,
                        y = NULL) +
      theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none") 
  
```



# Getting the top 10 Locations
```{r}
Top_10_Locations <- Location_count %>%
                     top_n(10, count) %>%
                  arrange(desc(count))
print(Top_10_Locations)
```






# Getting the top 10 percentages
```{r}
Top_10_Locations <- Top_10_Locations%>%
                     mutate(pct = count/ sum(count))
print(Top_10_Locations)
```


# Getting Lowest Locations
```{r}
Bottom_10_Locations <- Location_count %>%
                     top_n(-10, count) %>%
                    arrange(desc(count)) %>%
                    mutate(pct = count/sum(count))

print(Bottom_10_Locations)
```

# Top 10 Locations Bar Graphs For Crime

```{r}
ggplot(Top_10_Locations) + 
      geom_bar(mapping = aes(x = reorder(Location.Type,+pct),y = pct, fill = Location.Type), stat = 'identity') +
      coord_flip() + 
      labs(title = "Top 10 Locations By Incident 2015-2019",
                          x = NULL,
                          y = NULL) + 
  theme_classic() +
    theme(legend.position = "none")+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```




# Generating maps by crime- Registering Google key

```{r}
register_google(key = 'AIzaSyCwMV07Bb4MKO2fdUaVbJzIAXH_r5kyoq4')
```



# Subsetting for crime data for theft for 2018


```{r}
crimedatatheft <- cleandata
```


```{r}
crimedatatheft2018 <- cleandata %>%
                      filter(Occurred_Year == 2018,
                             Category.Description == 'Theft')
```

# Subsetting for crime data for theft for 2017

```{r}
crimedatatheft2017 <- cleandata %>%
                      filter(Occurred_Year == 2017,
                             Category.Description == 'Theft')
```

# Subsetting for crime data for theft for 2016

```{r}
crimedatatheft2016 <- cleandata %>%
                      filter(Occurred_Year == 2016,
                             Category.Description == 'Theft')
```


# # Subsetting for crime data for theft for 2015


```{r}
crimedatatheft2015 <- cleandata %>%
                      filter(Occurred_Year == 2015,
                             Category.Description == 'Theft')
```



# Making the coordinates for the maps


```{r}
coords <- cbind(Longitude = as.numeric(as.character(crimedatatheft$Longitude)), Latitude = as.numeric(as.character(crimedatatheft$Latitude)))
crime.pts <- SpatialPointsDataFrame(coords, crimedatatheft[, -(5:6)], proj4string = CRS("+init=epsg:4326"))
```


```{r}
coords <- cbind(Longitude = as.numeric(as.character(crimedatatheft2018$Longitude)), Latitude = as.numeric(as.character(crimedatatheft2018$Latitude)))
crime.pts <- SpatialPointsDataFrame(coords, crimedatatheft2018[, -(5:6)], proj4string = CRS("+init=epsg:4326"))
```



```{r}
coords <- cbind(Longitude = as.numeric(as.character(crimedatatheft2017$Longitude)), Latitude = as.numeric(as.character(crimedatatheft2017$Latitude)))
crime.pts <- SpatialPointsDataFrame(coords, crimedatatheft2017[, -(5:6)], proj4string = CRS("+init=epsg:4326"))
```

```{r}
coords <- cbind(Longitude = as.numeric(as.character(crimedatatheft2016$Longitude)), Latitude = as.numeric(as.character(crimedatatheft2016$Latitude)))
crime.pts <- SpatialPointsDataFrame(coords, crimedatatheft2016[, -(5:6)], proj4string = CRS("+init=epsg:4326"))
```

```{r}
coords <- cbind(Longitude = as.numeric(as.character(crimedatatheft2015$Longitude)), Latitude = as.numeric(as.character(crimedatatheft2015$Latitude)))
crime.pts <- SpatialPointsDataFrame(coords, crimedatatheft2015[, -(5:6)], proj4string = CRS("+init=epsg:4326"))
```

# Renaming Category.Description to Category

```{r}
names(crimedatatheft)[names(crimedatatheft) == 'Category.Description'] <- 'Category'
```


# Making geospatial maps

```{r}
#plot the  hybrid Google Maps basemap
map <- qmap('Austin', zoom =13, maptype = 'hybrid')
#plot the crime points on top
map + geom_point(data = crimedatatheft, aes(colour = Category,x = Longitude, y = Latitude), size=2, alpha=0.18
                 ) +
   labs(title = "Austin Crime - All Types (2015-2019)") + 
          theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ Occurred_Year, ncol = 3)
  
```


```{r}
#plot the  hybrid Google Maps basemap
map <- qmap('Austin', zoom = 12, maptype = 'hybrid')
#plot the crime points on top
map + geom_point(data = crimedatatheft2018, aes(x = Longitude, y = Latitude), color = 'pink',size=2, alpha=0.3) +
   labs(title = "Austin Crime Theft - 2018 By Month") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Occurred_Month, ncol = 6)
  
```

```{r}
#plot the  hybrid Google Maps basemap
map <- qmap('Austin', zoom = 12, maptype = 'hybrid')
#plot the crime points on top
map + geom_point(data = crimedatatheft2017, aes(x = Longitude, y = Latitude), color = 'pink',size=2, alpha=0.3) +
   labs(title = "Austin Crime Theft - 2017 By Month") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Occurred_Month, ncol = 6)
```  


```{r}
#plot the  hybrid Google Maps basemap
map <- qmap('Austin', zoom = 12, maptype = 'hybrid')
#plot the crime points on top
map + geom_point(data = crimedatatheft2016, aes(x = Longitude, y = Latitude), color = 'pink',size=2, alpha=0.3) +
   labs(title = "Austin Crime Theft - 2016 By Month") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Occurred_Month, ncol = 6)
``` 


```{r}
#plot the  hybrid Google Maps basemap
map <- qmap('Austin', zoom = 12, maptype = 'hybrid')
#plot the crime points on top
map + geom_point(data = crimedatatheft2015, aes(x = Longitude, y = Latitude), color = 'pink',size=2, alpha=0.3) +
   labs(title = "Austin Crime Theft - 2015 By Month") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Occurred_Month, ncol = 6)
``` 



